Usable Security in Virtual Power Plants

### 2018-01-23 acmart-master/draft.tex
Added more content

Critical infrastructures are integral to the functioning of our daily lives. They are changing in order to solve different problems it faces nowadays. The growth of power demand, integration of markets and the need of sustainable and efficient energy sources. \\
One such important critical infrastructure is the energy grid. A typical energy grid is divided into several domains such as generation, transmission and distribution. All the actors within domains of a energy grid are connected through information and communication technology(ICT). Hence, energy grids have continuously been the target of cyber attacks and provided their use of standard architecture and internet, they are relatively easier to attack. Security mechanisms have been introduced to protect against such attacks. Security-by-design approach has been used to determine security measures for various systems during design phase. With this approach, the system is equipped to deal with difficult scenarios before deployment. \\ However, there may be instances where the user must intervene to control the system when under attack and stress. During such instances, the user maybe asked to re-authenticate themselves, at which point remembering increasingly difficult passwords and entering correctly actually causes an overhead on the user. The security mechanism in fact is a hindrance here. Such a situation could be avoided if the authentication system in place is usable under different conditions.
%Intro big pic, narrow, Related work seperate subsection

@@

\subsection{Usability Standards: ISO 9241}

@@

ISO 9241-11 also emphasises that usability is dependent on the context of use and that the
level of usability achieved will depend on the specific circumstances in which a product is
used. The context of use consists of the users, tasks, equipment (hardware, software and
materials), and the physical and organisational environments which may all influence the
usability of a product (see \ref{fig:usability1}).

@@

\section{Framework for Usable Security}
Since, usability and security are normally at opposite ends of the spectrum, there are no uniform set of criteria or guidelines that define a usable security system. Nurse et al\cite{nurse2011guidelines} 
compiled a list as guidelines for usable security.
\newline
To design a security system that is usable, we need appropriate metrics to measure its efficiency. Therefore, we need to define such metrics. With the security system in focus, the usability metrics must stem from the properties of security themselves. Through this process, we then develop a security with a view of usability in the design stage itself, rather than first building a security system and then attempting to incorporate usability. When designing a security system, each component of security, every module could be tested for its usability wherever applicable. Usability of one component may affect that of another. That is, if a component is less usable, the overall usability of the system reduces, since the user may spend more resources on such a component at the cost of others. In other words, it hinders the performance of the system as a whole. The issue at hand is that due to the lack of usability metrics or any quantifiable value, any system cannot yet be defined as usable or unusable. \\

@@

Usability of a system, though determined by efficiency, effectiveness and satisfaction, is still not quantified. Furthermore, the three properties can be measured in various ways for different systems. To specialize usability for a security system, a relation between security and usability must be established. In the presence of such a relation, a security system's properties can be optimized, via metrics, such that usability properties are satisfied. This method can be used to integrate usability into any system instead of security systems. \\

In essence, each property has a metric with a desirable value that makes the system usable.
But these metrics are defined when a system operates in normal mode. When in any situation
other than normal mode, these metrics may vary and the system will have to adapt to the
situation. This variation could be used as an indication of how usable the system is.

A security system will have to go through a usability engineering cycle subject to constraints
and metrics. Ideally, security must not be an extra design layer, but it should be included in
the whole software design process of the systems used in grids. This may be particularly
hard to achieve, since smart grids at the moment use common operating systems(Windows,
linux), and therefore leaves this question open. It might be possible to develop these
common OSes with respect to the security required for smart grids, though this does not
appear to be considered as yet. Similarly, usability must be considered in the design phase as well. Individual components may be evaluated for usability. This may be considered as \textbf{Usable Security by Design.}

\subsection{Deriving Usability}
With respect to the smart grid, the asset we are trying to protect is the control and information flow of the grid. Therefore, being a critical infrastructure, their estimated value is evidently high.   
\newline The countermeasure, in this case, is an authentication system. To move forward with the framework, we need to identify the elements of the authentication system. Generally, they contain: An interface, a database, an authentication protocol run, key/token management and encryption algorithms. All the elements of the system can be measured using suitable metrics to quantify their impact on the usability of the system. In table \ref{tab:security-usability-relation}, different properties or factors affecting security are listed. The third column contains these factors, the second column contains the criteria for usability, which also represent the requirements for usability in this case and the first column contains the properties and factors affecting usability. The table is arranged such that if the security factors are optimized, they either directly satisfy the usability factors(or properties) in the first column or they satisfy the usability criteria(in the second column) and transitively satisfying the usability factors(or properties).

@@

Therefore, a \textit{usable security framework for an authentication system} consists of an authentication system which includes a user-friendly interface, requires a user to have minimalistic or easy to use authentication tokens, satisfies the criteria for a conventional security system such as Confidentiality, Integrity and Availability in addition to all the usability factors mentioned in \ref{tab:security-usability-relation}. The evaluation of these factors via usability criteria with respect to their metrics should fall in the permissible zone. The user also must provide feedback and specify the drawbacks, if any and suggestions to improve the system from the perspective of usability.

\subsection{Using The Framework}
To use the usable security framework, the organization intending to use it has to specify the goals on the system. Similar systems may be optimized in different ways, hence the developers need to specify the priorities. The important factors and properties affecting the system must be listed and with respect to the priorities, i.e, goals of the system and the developers, those factors affecting the user most should be selected. For example, let the goal be a highly efficient authentication mechanism. Therefore, accuracy is a goal and a priority.\\
The usability framework, with the help of QUIM, breaks down usability factors into criteria that can be quantified. Those usability factors that are of importance to the developers must be focused upon, however, the three usability factors, efficiency, effectiveness and satisfaction must all be used to satisfy usability, as per ISO 9241-11. In our example, accuracy of the system will be the focus and hence it will need to be optimized.\\
In the next step, those factors derived from the target system must be used as quantifiable values to measure the criteria specified in table \ref{tab:security-usability-relation}. If a criteria cannot be quantified, alternative criteria can be selected from the QUIM usability criteria. In any case, the prioritized factors must be quantified. In our example, accuracy is a measurable quantity of the system as well as the usability criteria used to determine the efficiency.\\
Accordingly, the factors measured from the system must be used to quantify the usability criteria, which help measure the usability factors.

@@

\subsection{Inference}
Now that we have results of two different authentication systems, a comparison can be made to assess which is more usable of the two. Tables \ref{tab:result1} and \ref{tab:result2} can be compared now that they have normalized metrics, i.e, common metrics. Since the EER is only used in biometric cases, it will not be used for comparison. Password based systems have a 100\% accuracy and a high rating for "Interface Design", but as explained earlier, these values were expected. Accuracy of biometric system observed as 96.1\% is still formidable. In case of a false negative, the task completion time, being exceptionally low, compensates for the time lost. The key factors are Task Completion Time, Interface Design and Memory Load. The fingerprint matching system used was exceptionally fast at 1.8 seconds for a successful authentication, although this can be attributed to relatively low sample space and the user of a capacitive scanner. In contrast a password system took an average of 12.2 seconds. This may be due to complexity of passwords, typing speed and additional time required to enter the user identifier such as a username. Interface design were both scored above 4, with password systems higher due to its familiarity to users. Memory Load, being a very important focus, due to difficulty in remembering difficult passwords especially under stress, shows that password systems are not preferable at all. Fingerprint systems require no memory use and are very usable.\\
It can be seen from the comparison that biometric systems such as fingerprint recognition and matching systems are accurate, faster, easy-to-use and are hence very usable.

\section{Example: Application}
To describe the approach for usability-by-design, or in this case usable security-by-design, the use case of DER scheduling in Dynamic Virtual Power Plants was used.

\begin{figure}[h]

@@

\def\unitfactor{0.5}
\newthread{a}{VPP SP}
\newinst[2]{b}{IED(all)}
%\newinst[2]{c}{Metering}
\newinst[2]{d}{DER}

@@

Using method in Uslar et. al\cite{uslar2014security}, the following mapping is obtained:

@@

@@

\section{Conclusion and Future Work}
Extensive security measures will be needed to protect power plants and other critical infrastructures from various kinds of attacks. In order to aid the operator in attack scenarios, the system in place should be usable. To assess usability of systems, a usability framework for security was derived, which helps establish the relation between the target system, in this case, security and usability.  With the use of this framework, it is possible to compare multiple systems and determine if they are usable or not. This was demonstrated by comparing two systems used in authentication mechanisms, a password based system and a fingerprint matching system. It is important to note that this framework can applied subjective to the system and the mechanism. Different systems will have different values for the framework. The framework paves the way to \textit{usable security by design}. Future Work will involve defining detailed metrics with methods or formulae to measure them. Metrics may also be defined to determine the degree of performance of a user under various scenarios.

>>>>

Critical infrastructures are integral to the functioning of our daily lives. They are changing in order to solve different problems it faces nowadays. The growth of power demand, integration of markets and the need of sustainable and efficient energy sources. \\
One such important critical infrastructure is the energy grid. A typical energy grid is divided into several domains such as generation, transmission and distribution. All the actors within domains of a energy grid are connected through information and communication technology(ICT). Hence, energy grids have continuously been the target of cyber attacks and provided their use of standard architecture and internet, they are relatively easier to attack. Security mechanisms have been introduced to protect against such attacks. Security-by-design approach has been used to determine security measures for various systems during design phase. With this approach, the system is equipped to deal with difficult scenarios before deployment. \\ However, there may be instances where the user must intervene to control the system when under attack and stress. During such instances, the user maybe asked to re-authenticate themselves, at which point remembering increasingly difficult passwords and entering it %correctly actually causes an overhead on the user. The security mechanism in fact is a hindrance here. Such a situation could be avoided if the authentication system in place is usable under different conditions.
%Intro big pic, narrow, Related work seperate subsection

@@

\subsection{Usability Standard: ISO 9241}

@@

ISO 9241-11 also emphasises that usability is dependent on the context of use and that the
level of usability achieved will depend on the specific circumstances in which a product is
used. The context of use consists of the users, tasks, equipment (hardware, software and
materials), and the physical and organisational environments which may all influence the
usability of a product (see figure \ref{fig:usability1}).

@@

\section{Framework for Usable Security}
In the previous section, usability was explained as a broad concept. It is applicable to any system which is directly interfaced by a user or has an impact on user perforamnce. In this section, we specialize the usability concept to security systems and as will further be explained, to authentication mechanisms.\\ Since, usability and security are normally at opposite ends of the spectrum, there are no uniform set of criteria or guidelines that define a usable security system. Nurse et al\cite{nurse2011guidelines} 
compiled a list as guidelines for usable security.
\newline
To design a security system that is usable, we need appropriate metrics to measure its efficiency. Therefore, we need to define such metrics. With the security system in focus, the usability metrics must stem from the properties of security themselves. Through this process, we then develop a security with a view of usability in the design stage itself, rather than first building a security system and then attempting to incorporate usability. When designing a security system, each component of security, every module could be tested for its usability wherever applicable. Usability of one component may affect that of another. That is, if a component is less usable, the overall usability of the system reduces, since the user may spend more resources on such a component at the cost of others. In other words, it hinders the performance of the system as a whole. The issue at hand is that due to the lack of usability metrics or any quantifiable value, any system cannot yet be defined as usable or unusable. \\

@@

Usability of a system, though determined by efficiency, effectiveness and satisfaction, is still not quantified. Furthermore, the three properties can be measured in various ways for different systems. To specialize usability for a security system, a relation between security and usability must be established. In the presence of such a relation, a security system's properties can be optimized, via metrics, such that usability properties are satisfied. \textit{This method can be used to integrate usability into any system instead of security systems.} \\

In essence, each property has a metric with a desirable value that makes the system usable. But these metrics are defined when a system operates in normal mode. When in any situation other than normal mode, these metrics may vary and the system will have to adapt to the situation. This variation could be used as an indication of how usable the system is. A security system will have to go through a usability engineering cycle subject to constraints and metrics. To achieve strong security, security engineering must begin parallel to system development, which is known as security-by-design. Similarly, usability must be considered in the design phase as well. Individual components may be evaluated for usability. This may be considered as \textbf{Usable Security by Design.}

\subsection{Deriving Usability for Authentication}
With respect to the virtual power plants, the asset we are trying to protect is the control and information flow of the grid. Therefore, being a critical infrastructure, their estimated value is evidently high.   
\newline The countermeasure, in this case, is an authentication system. To move forward with the framework, we need to identify the elements of the authentication system. Generally, they contain: An interface, a database, an authentication protocol run, key/token management and encryption algorithms. All the elements of the system can be measured using suitable metrics to quantify their impact on the usability of the system. In table \ref{tab:security-usability-relation}, different properties or factors affecting security are listed. The third column contains these factors, the second column contains the criteria for usability, which also represent the requirements for usability in this case and the first column contains the properties and factors affecting usability. The table is arranged such that if the security factors are optimized, they either directly satisfy the usability factors(or properties) in the first column or they satisfy the usability criteria(in the second column) and transitively satisfying the usability factors(or properties).

@@

Therefore, a \textit{usable security framework for an authentication system} consists of an authentication system which includes a user-friendly interface, requires a user to have minimalistic or easy to use authentication tokens, satisfies the criteria for a conventional security system such as Confidentiality, Integrity and Availability in addition to all the usability factors mentioned in \ref{tab:security-usability-relation}. The evaluation of these factors via usability criteria with respect to their metrics should fall in the permissible zone. The user also must provide feedback and specify the drawbacks, if any and suggestions to improve the system from the perspective of usability. \\
The factors mentioned in the third column are specific to authentication mechanisms. If usability was being considered for any other system, the respective factors would be derived and placed accordingly in that column. For example, consider an organization attempting to develop self-driving automobiles. User experience plays a big part in such an inventive technology. To measure usability of such vehicles, effectiveness will be determined by the performance of the vehicle. It is up to the discretion of the organization to determine which metric they would prefer to use to measure the performance. An example could be speed or acceleration. Efficiency can be measured by fuel consumption. Satisfaction can be measured by user feedback. In this way, the framework could be used for any system.

\subsection{Using The Framework}
To use the usable security framework, the organization intending to use it has to specify the goals on the system. Similar systems may be optimized in different ways, hence the developers need to specify their priorities. The important factors and properties affecting the system must be listed and with respect to the priorities, i.e, goals of the system and the developers, those factors affecting the user most should be selected. For example, let the goal be a highly user-friendly yet efficient authentication mechanism. Therefore, Interface Design and Time as a resource are goals and are priorities.\\
The usability framework, with the help of QUIM, breaks down usability factors into criteria that can be quantified. Those usability factors that are of importance to the developers must be focused upon, however, the three usability factors, efficiency, effectiveness and satisfaction must all be used to satisfy usability, as per ISO 9241-11. In our example, task completion time and memory load will be the focus and hence it will need to be optimized.\\
In the next step, those factors derived from the target system must be used as quantifiable values to measure the criteria specified in table \ref{tab:security-usability-relation}. If a criteria cannot be quantified, alternative criteria can be selected from the QUIM usability criteria. In any case, the prioritized factors must be quantified. In our example, accuracy is a measurable quantity of the system as well as the usability criteria used to determine the efficiency.\\
Accordingly, the factors measured from the system must be used to quantify the usability criteria, which help measure the usability factors.

@@

\subsection{Inference}
Now that we have results of two different authentication systems, a comparison can be made to assess which is more usable of the two. Tables \ref{tab:result1} and \ref{tab:result2} can be compared now that they have normalized metrics, i.e, common metrics. Since the EER is only used in biometric cases, it will not be used for comparison. Password based systems have a 100\% accuracy and a high rating for "Interface Design", but as explained earlier, these values were expected. Accuracy of biometric system observed as 96.1\% is still formidable. In case of a false negative, the task completion time, being exceptionally low, compensates for the time lost. The key factors are Task Completion Time, Interface Design and Memory Load. The fingerprint matching system used was exceptionally fast at 1.8 seconds for a successful authentication, although this can be attributed to relatively low sample space and the use of a capacitive scanner. In contrast a password system took an average of 12.2 seconds. This may be due to complexity of passwords, typing speed and additional time required to enter the user identifier such as a username. Interface design were both scored above 4, with password systems higher due to its familiarity to users. Memory Load, being a very important focus, due to difficulty in remembering difficult passwords especially under stress, shows that password systems are not preferable at all. Fingerprint systems require no memory use and are very usable.\\
It can be seen from the comparison that biometric systems such as fingerprint recognition and matching systems are accurate, faster, easy-to-use and are hence very usable.\\
A conclusion drawn from the analysis is that there can be no standard values for a usable system, i.e, one set of values for the usability factors cannot be applied to all systems. Usability values are subjective to the system and to the goals of the organization. It could be possible to set a threshold for a specific type of system. For example, several single factor authentication systems could be tested. The system that is determined to be the most usable may have its values set as a threshold for usable authentication mechanisms. However, these values of the framework can only be applied to single factor authentication mechanisms and no other system. This could be extended to two factor authentication mechanisms and other systems as well.

\section{Example Use Case}
To describe the approach for usability-by-design, or in this case usable security-by-design, the use case of DER scheduling in Dynamic Virtual Power Plants was used. The figure \ref{fig:uml_diag} depicts a UML diagram of a typical scheduling process that takes place in Dynamic Virtual Power Plants\cite{niesse2014conjoint}. The distributed agents in the field are responsible for scheduling based on information such as predicted energy generation from its Distributed Energy Resource(DER), Market information, available agents in the area etc. and forms a coalition of energy sources to meet the energy demand.

\begin{figure}[H]

@@

\def\unitfactor{0.4}
\newthread{a}{VPP SP}
\newinst[2]{b}{IED(all)}
%\newinst[2]{c}{Metering}
\newinst[2]{d}{DER}

@@

Employing the method in Uslar et. al\cite{uslar2014security}, the mapping shown in figure \ref{fig:HLdiag} is obtained using the NIST 7628 logical reference diagram\cite{NIST7628:2014}. There are 4 identified actors and their respective logical interfaces. The actor "1 Plant Control System" corresponds to the DER at the generation side, "46 Transmission IED" is the field agent on an Intelligent Electronic Device(IED) as described in Dynamic Virtual Power Plants, "30 Energy Management System" is an actor within the operation domain which corresponds to the Virtual Power Plant Service Provider interface between the system and the IED. All communication takes place through the "37 Transmission SCADA". Each actor is interfaced with Transmission SCADA as it is the actor that represents the transmission in a power system. Each interface has a set of specific security requirements described in NIST 7628.

@@

Since there may be several interfaces with common requirements, NIST 7628 categorizes interfaces into categories with similar requirements. For example, interface U83 falls under category 1. Under category 1, security requirements such as AC-14, IA-4, IA-6 and so on are specified. AC-14, for example, is a requirement that allows certain actions, under certain circumstance such as emergencies, to be performed without identification and authentication of the operator. Such a requirement can actually be exploited by a malicious outsider or an attacker. An attacker may simply induce an attack or a false alarm, in which case security mechanisms such as authentication procedures will be suspended. The attacker then has free access to the system with the authentication procedure bypassed. It is understandable that such a requirement is recommended, in order to prevent the operator from being locked out of the system during emergencies. However, it clearly has flaws which can be exploited. A suitable solution for this flaw is the introduction of an usable authentication mechanism that secures the system access as well as reduces the burden of authentication on the operator by its usability.\\

A usable authentication mechanism can be developed to aid the operators of power plants. Developing such a mechanism requires user preferences prior to its implementation. A survey can be conducted to obtain the opinions and preferences of grid operators. This survey will have to conducted early, i.e, prior to the requirement specification. More detail on the method to integrate usability into the software development lifecycle is described in the following subsection.

@@

\section{Conclusion and Future Work}
Extensive security measures will be needed to protect power plants and other critical infrastructures from various kinds of attacks. In order to aid the operator in attack scenarios, the system in place should be usable. To assess usability of systems, a usability framework for security was derived, which helps establish the relation between the target system, in this case, security and usability.  With the use of this framework, it is possible to compare multiple systems and determine if they are usable or not. This was demonstrated by comparing two systems used in authentication mechanisms, a password based system and a fingerprint matching system. It is important to note that this framework can applied subjective to the system and the mechanism. Different systems will have different values for the framework and different systems factors may be related to usability criteria and factors. The framework paves the way to \textit{usable security by design}. Future Work will involve defining detailed metrics with methods or formulae to measure them. Metrics may also be defined to determine the degree of performance of a user under various scenarios. Specific usability frameworks for various systems may also be derived.
